---
title: "GameDataset DATA WRANGLING"
author: "VEDANT SINGH JADON 22BDS0101"
date: "2025-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

This R Markdown document processes and analyzes the Game dataset. It includes:
- Data cleaning
- Feature engineering
- Exploratory data analysis (EDA)
- Visualizations

---

## 2. Load Necessary Libraries

```{r}
# Load essential libraries
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(knitr)
```

---

# 3. Data Wrangling

## 3.1 Load and Explore Data

```{r}
# Reading data with proper NA handling
games <- read.csv("Game_dataset_testing.csv", na.strings = c("NA", "NULL"))

# Initial examination
dim(games) # Check dimensions
str(games) # View structure
summary(games) # Basic stats
glimpse(games)

# Checking missing values
colSums(is.na(games))
```

---

## 3.2 Data Cleaning

```{r}
# Converting date columns to proper format
games <- games %>%
  mutate(
    Release.date = as.Date(Release.date, format = "%B %d, %Y"),
    released = as.Date(released)
  )

# Cleaning 'Total copies sold'
games <- games %>%
  mutate(
    copies_millions = as.numeric(str_extract(Total.copies.sold, "\\d+\\.?\\d*")),
    Total.copies = copies_millions * 1000000
  ) %>%
  select(-Total.copies.sold, -copies_millions)

# Handling date discrepancies
games <- games %>%
  mutate(
    Release.date = coalesce(Release.date, released),
    release_year = year(Release.date)
  ) %>%
  select(-released)

# Remove redundant columns
games <- games %>% select(-platforms)

# Handle missing values
games <- games %>%
  mutate(
    metacritic = ifelse(metacritic < 0 | metacritic > 100, NA, metacritic)
  )

# Remove duplicates
games <- games[!duplicated(games$Game), ]

# Handle missing values in Categorical column
games <- games%>%
  mutate(Series = ifelse(Series == "—" | is.na(Series), "Standalone", Series))
```

---

## 3.3 Feature Engineering

```{r}
# Creating series indicator
games <- games %>%
  mutate(
    is_series = ifelse(Series == "—", "No", "Yes")
  )

# Creating rating categories
games <- games %>%
  mutate(
    rating_category = case_when(
      rating >= 4.5 ~ "Excellent",
      rating >= 4.0 ~ "Good",
      rating >= 3.0 ~ "Average",
      TRUE ~ "Below Average"
    )
  )
```

---

# 4. Data Validation

```{r}
# Final quality checks
glimpse(games)
summary(games)

# Check date ranges
ggplot(games, aes(x = release_year)) +
  geom_histogram(binwidth = 1) +
  ggtitle("Release Year Distribution")

# Validate numeric ranges
ggplot(games, aes(x = metacritic)) +
  geom_histogram(bins = 20) +
  ggtitle("Metacritic Score Distribution")

summary(games$Total.copies)
```

---

# 5. Export Cleaned Data

```{r}
write.csv(games,"GameSet_Cleaned.csv",row.names = FALSE)
```

---

# 6. Data Visualization

## 6.1 Table Overview

```{r}
kable(
  head(games, 10), 
  align = c("l", "l", "l", "l", "l", "l", "c", "c", "r", "c", "c", "l"),
  col.names = c(
    "Game", "Series", "Release Date", "Genre(s)", "Developer(s)", 
    "Publisher(s)", "Rating", "Metacritic", "Total Copies", 
    "Release Year", "In Series?", "Rating Category"
  )
)

summary_df<-summary(games)
kable(summary_df,caption = "Summary of PC Games Dataset")


```

---

## 6.2 Sales Analysis

```{r}
# Top 10 best-selling games
top_sellers <- games %>%
  arrange(desc(Total.copies)) %>%
  head(10)

ggplot(top_sellers, aes(x = reorder(Game, Total.copies), y = Total.copies/1e6)) +
  geom_col(fill = "steelblue") +
  labs(title = "Top 10 Best-Selling Games",
       x = "Game", y = "Copies Sold (Millions)") +
  coord_flip()
```

---

## 6.3 Temporal Trends

```{r}
# Sales by release year
yearly_sales <- games %>%
  group_by(release_year) %>%
  summarise(Total_Sales = sum(Total.copies, na.rm = TRUE))

ggplot(yearly_sales, aes(x = release_year, y = Total_Sales/1e6)) +
  geom_line(color = "darkgreen", linewidth = 1) +
  geom_point(color = "darkgreen") +
  labs(title = "Game Sales Trends Over Time",
       x = "Release Year", y = "Total Copies Sold (Millions)")
```

---

## 6.4 Rating Analysis

```{r}
# Rating vs Sales
ggplot(games, aes(x = rating, y = Total.copies/1e6)) +
  geom_smooth(method = "loess", color = "darkgreen", fill = "lightgreen") +
  labs(title = "Trend of Ratings vs Sales",
       x = "User Rating", y = "Copies Sold (Millions)") +
  theme_light()
```

```{r}
# Rating distribution by series status
ggplot(games, aes(x = is_series, y = rating, fill = is_series)) +
  geom_boxplot() +
  labs(title = "Rating Distribution: Series vs Standalone Games",
       x = "Part of Series", y = "Rating")
```

---

## 6.5 Genre Analysis

```{r}
# Most common genres (top 10)
genre_counts <- games %>%
  count(Genre.s.) %>%
  arrange(desc(n)) %>%
  head(10)

ggplot(genre_counts, aes(x = reorder(Genre.s., n), y = n)) +
  geom_col(fill = "orange") +
  labs(title = "Most Common Game Genres",
       x = "Genre", y = "Count") +
  coord_flip()
